---
title: "The Gravity Model"
author: "Jesus Davila- Sara Savino- Adoracion Hernandez"
date: "10/2021"
output:
  pdf_document: default
  html_document: default
---

## Setup

```{r packages, include=FALSE}
pacman::p_load(here, rjson, dplyr, ggplot2, gridExtra, kableExtra,reshape2,lubridate,tidyr,tidyverse, estimatr, foreign, stats, psych, stargazer,texreg, car, vtable)
```

```{r, include=FALSE}
here::here()
```

```{r setup, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
bilateralData <- read.dta("servicesdataset 2.dta")
TSPdata <- bilateralData[bilateralData$sector=="TSP",]
```

## Data Cleaning and Correlation

```{r sum, warning=FALSE}
#Prepare data for correlation-4 variables only 
TSPdataGRM <- TSPdata[,c("trade","dist","gdp_exp","gdp_imp")]
#filter out all rows containing 0 
row_sub = apply(TSPdataGRM, 1, function(row) all(row !=0 ))
TSPdataGRM <- TSPdataGRM[row_sub,]
#prepare correlation and remove all NA or zero values 
correlation <- cor(log(TSPdataGRM), use="na.or.complete")
correlation

```

```{r sum2, warning=FALSE}
#QUESTION 1
#Prepare data for descriptive statistics for following variables: trade,  distance,  GDP  of  importer  and  exporter, contiguity,  common  language,  common  colonizer  and  colony

TSPdataDS <- TSPdata[,c("trade","dist","gdp_exp","gdp_imp", "contig", "comlang_off", "colony", "comcol")]

# Filter all rows where trade = 0 
TSPdataDS <- filter_at(TSPdataDS, vars("trade"), all_vars((.) != 0))
sumtable(TSPdataDS)
dim(TSPdataDS)

#stargazer(TSPdataDS, type = "html")

```


## 

### Scatter plot and line of best fit for trade versus distance.

```{r graph1, warning=FALSE}
# CODE FOR QUESTION 2
# TRADE AND DISTANCE
df = read.dta("servicesdataset 2.dta", convert.f=TRUE)
logdistance <- log(TSPdataGRM$dist)
logtrade <- log(TSPdataGRM$trade)
plot1 <- ggplot(TSPdataDS, aes(x = logdistance, y = trade,)) +
 geom_point(shape=1) + 
 geom_smooth(method=lm, # Add linear regression line
 se=FALSE) + # Don't add shaded confidence region
 theme_classic() + #no background
 ggtitle("Trade versus Distance combined") +
 theme(plot.title = element_text(hjust = 0.5)) + # put title in the middle
 labs(y="Trade", x = "Log(Distance)") + # set names for axes
 scale_y_log10(labels = scales::comma)# ln base scale
plot1
```

### Scatter plot and line of best fit for trade versus combined GDP. 

```{r graph2, warning=FALSE}
#TRADE AND COMBINED GDP
logdifference <- log(TSPdataGRM$gdp_exp * TSPdataGRM$gdp_imp)
plot2 <- ggplot(TSPdataDS, aes(x = logdifference, y = trade,)) +
 geom_point(shape=1) + 
 geom_smooth(method=lm, # Add linear regression line
 se=FALSE) + # Don't add shaded confidence region
 theme_classic() + #no background
 ggtitle("Trade versus Log GDP combined") +
 theme(plot.title = element_text(hjust = 0.5)) + # put title in the middle
 labs(y="Trade", x = "log(GDP combined)") + # set names for axes
 scale_y_log10(labels = scales::comma)# ln base scale
plot2
```

### Scatter plot and line of best fit for trade versus combined GDP (improved) - clustered

```{r graph3, warning=FALSE}
#TRADE DISTANCE CLUSTER CONTIGUITY
plot3 <- ggplot(TSPdataDS, aes(x=logdistance,
                         y=trade, colour=factor(contig))) +
                        geom_point(shape=1) + 
                        geom_smooth(method=lm, # Add linear regression line
                                se=FALSE,   # Don't add shaded confidence region
                                ) + 
                        theme_classic() + #no background
                        ggtitle("Log Distance vs Trade (Clustered by Contiguity)") +
                        theme(plot.title = element_text(hjust = 0.5)) + # put title in the middle
                        labs(y="Trade", x = "Log(Distance)") + # set names for axes
                        scale_y_log10(labels = scales::comma)# ln base scale
plot3
```
```{r graph4, warning=FALSE}
# COMBINED GDP TRADE CONTINUITY
plot4 <- ggplot(TSPdataDS, aes(x=log(gdp_exp * gdp_imp),
                         y=trade, colour=factor(contig))) +
                        geom_point(shape=1) + # Use hollow circles
                        geom_smooth(method=lm, # Add linear regression line
                                se=FALSE,   # Don't add shaded confidence region
                                ) + 
                        theme_classic() + #no background
                        ggtitle("Trade versus GDP (Clustered by Contiguity)") +
                        theme(plot.title = element_text(hjust = 0.5)) + # put title in the middle
                        labs(y="Trade", x = "log(GDP Combined)") + # set names for axes
                        scale_y_log10(labels = scales::comma)# ln base scale
plot4
```

```{r graph5, warning=FALSE}
# COMBINED TRADE DISTANCE LANGUAGE
plot5 <- ggplot(TSPdataDS, aes(x=logdistance,
                         y=trade, colour=factor(comlang_off))) +
                        geom_point(shape=1) + # Use hollow circles
                        geom_smooth(method=lm, # Add linear regression line
                                se=FALSE,   # Don't add shaded confidence region
                                ) + 
                        theme_classic() + #no background
                        ggtitle("Log Distance vs Trade comm. lan.") +
                        theme(plot.title = element_text(hjust = 0.5)) + # put title in the middle
                        labs(y="Trade", x = "Distance") + # set names for axes
                        scale_y_log10(labels = scales::comma)# ln base scale
plot5

```
```{r graph6, warning=FALSE}
# COMBINED GDP TRADE COMMON LANGUAGE
plot6 <- ggplot(TSPdataDS, aes(x=log(gdp_exp * gdp_imp),
                         y=trade, colour=factor(comlang_off))) +
                        geom_point(shape=1) + # Use hollow circles
                        geom_smooth(method=lm, # Add linear regression line
                                se=FALSE,   # Don't add shaded confidence region
                                ) + 
                        theme_classic() + #no background
                        ggtitle("Trade versus GDP common language") +
                        theme(plot.title = element_text(hjust = 0.5)) + # put title in the middle
                        labs(y="Log(Trade)", x = "log(GDP combined)") + # set names for axes
                        scale_y_log10(labels = scales::comma)# ln base scale
plot6
```

## Estimating the Intuitive Gravity Model

A command that performs OLS with robust clustered standard errors and equivalent to the regress command in Stata through, is lm_robust from the estimatr package. It takes the following format:

```{r reg1, warning=FALSE}
#QUESTION 3
TSPdataLR <- TSPdata[complete.cases(TSPdata[ ,
c("dist","gdp_exp","gdp_imp")]),]
TSPdataLR <- TSPdataLR[TSPdataLR$trade != 0, ]
reg1 <- lm_robust(log(trade) ~ log(dist) + log(gdp_imp) + log(gdp_exp) +
contig + comlang_off + colony + comcol, cluster = dist, data = TSPdataLR,
se_type = "stata")
summary(reg1)
```
### Hypothesis tests for both GDP coefficients are equal to unity and all historical and cultural coefficients are equal to zero.
#Question 3.5 
Testing a compound hypothesis in R can be done through the linearHypothesis command in the car package. Since we are using robust clustered standard errors, we have to adjust for the degrees of freedom. We can extract the new p-value by using the pf command and specifying our degrees of freedom. (In this case they are the number of coefficients we are testing for and the number of clusters minus one.)

```{r HT1, warning=FALSE}
# prepare the data as required in Box 1 and Box 7, we create new
# variables in order to test linear hypotheses
TSPdataLR$ln_gdp_imp <- log(TSPdataLR$gdp_imp)
TSPdataLR$ln_gdp_exp <- log(TSPdataLR$gdp_exp)
reg1_star <- lm_robust(log(trade) ~ log(dist) + ln_gdp_imp + ln_gdp_exp +
contig + comlang_off + colony + comcol,
cluster = dist, data = TSPdataLR, se_type = "stata")
```


```{r HT2, warning=FALSE}
ftest1 <- linearHypothesis(reg1_star,c("ln_gdp_imp = 1",
"ln_gdp_exp = 1"), test = "F")
ftest1

pf(ftest1$F, 2, length(unique(TSPdataLR$dist))-1, lower.tail=FALSE)
```

```{r HT, warning=FALSE}
ftest2 <- linearHypothesis(reg1_star,c("contig", "comlang_off", "colony",
"comcol"), test = "F")

ftest2

pf(ftest2$F, 4, length(unique(TSPdataLR$dist))-1, lower.tail=FALSE)
```


### OLS estimates of an augmented gravity model.

```{r reg2, warning=FALSE}
# prepare the data as required in Box 1 and Box 7
reg2 <- lm_robust(log(trade) ~ log(dist) + log(gdp_imp) + log(gdp_exp) +
contig + comlang_off + colony + comcol + etcr_exp +
etcr_imp, cluster = dist, data = TSPdataLR,
se_type = "stata")
summary(reg2)


# 1 coefficient



```



## Estimating the Structural Gravity Model

### OLS estimates of a gravity model with fixed effects by importer and exporter.

```{r reg3}
#QUESTION 4
TSPdataFX1 <- TSPdata[TSPdata$trade!=0 & TSPdata$dist!=0,]
reg3 <- lm_robust(log(trade) ~ log(dist) + contig + comlang_off +
colony + comcol + exp + imp,
cluster = dist, data = TSPdataFX1, se_type = "stata")
summary(reg3)
```
```{r reg4}
#QUESTION 5
# make sure you have the data loaded and transformed as in Box 1
# include only complete cases for variables for which we take log
OECDdummy <- TSPdata %>%
  mutate(oecd = ifelse(is.na(TSPdata$etcr_exp)|is.na(TSPdata$etcr_imp), 0, 1))

OECDdumDS <- OECDdummy[,c("trade", "dist", "gdp_exp","gdp_imp","contig", "comlang_off", "colony", "comcol", "oecd")]



OECDLR <- OECDdumDS[complete.cases(OECDdumDS[ ,
c("dist","gdp_exp","gdp_imp")]),]

OECDLR <- OECDdumDS[OECDdumDS$trade != 0, ]


reg4 <- lm_robust(log(trade) ~ log(dist) + log(gdp_imp) + log(gdp_exp) + contig + comlang_off + colony + comcol + oecd, cluster = dist, data = OECDLR,
se_type = "stata")
summary(regoecd)

```




## Summary of Results

```{r models, results = 'asis', warning=FALSE}

# load package
#library("dplyr")
#tab_model(reg1)

#stargazer(reg1a, type = 'html')
#stargazer(TSPdataDS, type = "text")

# texreg::screenreg(list(reg1, reg2, reg3),     # if knitting to pdf
#                   omit.coef=c('exp|imp'),
# #custom.coef.names = reglabels,
# include.ci = FALSE,
# caption = '',
# custom.note = 'Note: robust standard errors.')

texreg(list(reg1, reg3, reg4),     # if knitting to pdf
                  omit.coef=c('exp|imp'),
#custom.coef.names = reglabels,
include.ci = FALSE,
caption = '',
custom.note = 'Note: robust standard errors.')

# htmlreg(list(reg1, reg2, reg3),  # if knitting to html
# #custom.coef.names = reglabels,
# include.ci = FALSE,
# caption = '',
# custom.note = 'Note: robust standard errors.')

```

